%%
%% VERSION HISTORY
%%    22 May 2006 - John Papandriopoulos - Original version
%%    12 Jul 2007 - John Papandriopoulos - Converted into template
%%

\begin{acknowledgements}
First and foremost, I would like to express my sincere appreciation to my supervisor Prof. Karin Verspoor for the guidance despite of her heavy workload plus other academic duties. She provided valuable suggestions and insightful feedback for my thesis. I am truly thankful to have such a friendly, humorous and knowledgeable Professor as my supervisor.\\
I am grateful to Loi Nguyen, Thao Phuong Nguyen. I would have been in a much more miserable state if not for their support during my time in Australia. I am also lucky to have a chance to collaborate with and learn so much from my postgraduate classmates Tri Chandra, Steven Spratley, Victor Dilks and Harvey Nguyen.\\
I would like to thank Nhat-Duc Nguyen, Hung Viet Cao, Son-Tung Hoang, Ha Bach, Minh Duc Tran and other undergraduate classmates. We had so much fun in college and they encouraged me to take on further study.\\
My affectionate thank goes to my relatives Thuy-Trang Nguyen, Van-Trang Ho, Lan-Huong Hoang, Duc Nguyen, Bich Nguyen, Nam-Huan Hoang, Van-Anh Nguyen for the emotional and esteem support.\\
I am greatly indebted to Duy-Khiem Nguyen, Tuan Pham, Binh Pham Van, mentors, staff and other colleagues at NashTech Vietnam who have taught me not only domain knowledge but also professional attitude in the first days of my career.\\
I would like to extend my thank to friends who have shared various parts of my life: Anh-Tu Tran, Quang-Duy Le, Son-Vu Le, Ngoc-Hiep Tran, Tri-Khang Le, Thanh-Dat Nguyen, Dung Tran and many many others to whom I am thankful. I feel guilty that I cannot list all their names here. They have made my life much more enjoyable than I could ever hope for.\\
Finally, I wish to express my heartfelt gratitude, love and deepest appreciation to my family. My parents, Boi-Huong Hoang and Hien Ho Dac, always support my choice even though it is a difficult path. My elder sister Quynh-Nga Ho Dac always warm my heart with her cheerful attitude and I appreciate the consideration from my brother-in-law Manh-Ha Nguyen. They have given me unconditional love and it became my motivation. Without them, I would have never made this far.\\

\end{acknowledgements}
